A Partitioning Algorithm for Maximum Common Subgraphs\thanks{This work was supported by the Engineering and Physical Sciences Research Council [grant numbers EP/K503058/1 and EP/M508056/1]}

### 2017-02-15 paper/paper.tex
More introduction, and some tidying

To determine the similarity or difference between two graphs, we must first
find what they have in common
\citep{DBLP:journals/prl/Bunke97,DBLP:journals/prl/FernandezV01,KriegeThesis}.
The \emph{maximum common subgraph} family of problems involves finding a large
graph which is isomorphic to subgraphs of two given graphs simultaneously.
Because graphs are widely used to model real-world phenomena, maximum common
subgraph problems have arisen in ...  The problems are NP-hard, and remain
challenging computationally. Recent practical progress has been made by using
constraint programming
\citep{DBLP:conf/cp/NdiayeS11,DBLP:conf/cp/McCreeshNPS16}, by reducing to the
maximum clique problem \citep{LeviG,DBLP:conf/cp/McCreeshNPS16}, and
by adapting subgraph isomorphism algorithms \citep{UpcomingAAAIPaper}. This paper
introduces a new branch and bound algorithm which exploits special properties
of these problems to allow a much faster exploration of the search space,
whilst retaining the filtering and bounding benefits of the constraint
programming approach. We describe the algorithm for the basic maximum common
subgraph problem, and discuss how it may be adapted to handle vertex labels,
edge labels, and the requirement that the found subgraph be connected. We then
present an empirical study of the algorithm, demonstrating that it improves the
state of the art by more than an order of magnitude on the unlabelled variant
of the problem, and showing that it can handle much larger instances than
earlier constraint programming or clique approaches due to lower memory usage.

@@

\begin{algorithm}
\DontPrintSemicolon
\nl $\FuncSty{expand}(\AlgVar{future},M)$ \;
\nl \Begin{
%\nl \lIf {$\AlgVar{future} = \emptyset$ \bf{and} $|M| > |\AlgVar{incumbent}|$}
\nl \lIf {$|M| > |\AlgVar{incumbent}|$}{$save(matching)$} \label{StoreIncumbent}
%\nl \lIf {$\AlgVar{future} = \emptyset$}{return}
\nl $bf \gets 0$ \label{SetBestFutureToZero} \;
\nl \lFor {$\langle P,T \rangle \in \AlgVar{future}$}{$\AlgVar{bf} \gets \AlgVar{bf} + \min(|P|,|T|)$}
\nl \lIf {$|M|  + \AlgVar{bf} \leq |\AlgVar{incumbent}|$}{return} \label{PruneSearch}
\bigskip
\nl $\langle P,T \rangle \gets \FuncSty{SelectLabelClass}(\AlgVar{future})$ \label{SelectClass} \;
\nl $v \gets \FuncSty{SelectVertex}(P)$ \label{SelectVertex} \;
\nl \For {$w \in T$ \label{WLoop}} {
\nl    $M \gets M \cup (v,w)$ \label{GrowM} \;
\nl    \bf{Let} $\AlgVar{future'} \gets \emptyset$ \label{NewFuture} \;
\nl    \For {$\langle P',T'\rangle \in future$ \label{InnerLoop}}{
\nl        \bf{Let} $P'' \gets P' \cap N(v,G) \setminus \{v\}$ \label{NewPWithEdge} \;
\nl        \bf{Let} $T'' \gets T' \cap N(w,H) \setminus \{w\}$ \;
\nl        \If {$P'' \neq \emptyset$ \bf{and} $T'' \neq \emptyset$}
        {$\AlgVar{future'} \gets \AlgVar{future'} + \langle P'' , T'' \rangle$ \label{AddToFutureWithEdge}}
\nl        \bf{Let} $P'' \gets P' \cap N(v,\overline{G}) \setminus \{v\}$ \label{NewPWithoutEdge}  \;
\nl        \bf{Let} $T'' \gets T' \cap N(w,\overline{H}) \setminus \{w\}$ \;
\nl        \If {$P'' \neq \emptyset$ \bf{and} $T'' \neq \emptyset$}
        {$\AlgVar{future'} \gets \AlgVar{future'} + \langle P'' , T'' \rangle$} \label{InnerLoopEnd}
       }
\nl   $\FuncSty{expand}(\AlgVar{future'},M)$ \label{ExpandWithV} \;
\nl   $M \gets M \setminus \{(v,w)\}$ \label{ShrinkM} \;
  }
\nl $P' \gets P \setminus \{v\}$ \label{RemoveV} \;
\nl $\AlgVar{future} \gets \AlgVar{future} \setminus \{\langle P,T \rangle\}$\;
\nl \lIf {$P' \neq \emptyset$} {$\AlgVar{future} \gets \AlgVar{future} \cup \{\langle P',T \rangle \}$}
\nl $\FuncSty{expand}(\AlgVar{future},M)$ \label{ExpandWithoutV} \;
}
\;
\nl $\FuncSty{McSplit}(G,H)$ \label{McSplitFun} \;
\nl \Begin{
\nl $\AlgVar{incumbent} \gets \emptyset$ \;
\nl $\FuncSty{expand}(\{\langle V(G),V(H) \rangle \},\emptyset)$ \label{FirstExpandCall} \;
\nl return $|\AlgVar{incubent}|$ \;
}
\caption{\McSplit\ algorithm for Maximum Common Subgraph}
\label{jtAlg}
\end{algorithm}

@@

\begin{figure}
    \centering
    \subfigure[][MCS unlabelled undirected instances] {
        \centering
        \includegraphics*{gen-graph-plain-cumulative.pdf}
        \label{figure:plain-cumulative}
    }

@@

\paragraph{Large subgraph isomorphism instances} We also ran the algorithms on
a set of 5725 larger instances used in recent studies of subgraph
isomorphism~\cite{thelionpaper} and maximum common
subgraph~\cite{UpcomingAAAIPaper}.  This benchmark set includes real-world
graphs, along with graphs generated using several random models.  Pattern
graphs range from 4 vertices to 900 with a median of 80; target graphs range
from 10 vertices to 6671 with a median of 561.  Cumulative runtimes including
top-down algorithm from~\cite{UpcomingAAAIPaper} which was the previous state
of the art for this benchmark, are in \cref{figure:sip-cumulative}.  This is a
challenging set of instances, and more than half of the instances cannot be
solved within a timeout of 1000 seconds by any of the solvers. \McSplit\ is the
strongest solver overall for these instances.

>>>>

To determine the similarity or difference between two graphs, we must first
find what they have in common
\citep{DBLP:journals/prl/Bunke97,DBLP:journals/prl/FernandezV01,KriegeThesis}.
The \emph{maximum common subgraph} family of problems involves finding a large
graph which is isomorphic to subgraphs of two given graphs simultaneously.
Because graphs are widely used to model real-world phenomena, maximum common
subgraph problems have arisen in molecular science (where graphs represent
molecules)
\citep{DBLP:journals/jcamd/RaymondW02a,Ehrlich:2011,DAM2014,Grindley1993707},
computer vision \cite{DBLP:journals/jair/CookH94}, discovery of repeated
patterns in source code \cite{DBLP:journals/tkde/DjokoCH97} and circuits
\cite{DBLP:journals/jair/CookH94}, and recognition of handwritten characters
\cite{SIWEILU1991617}.

Maximum common subgraph problems are NP-hard, and remain challenging
computationally. Recent practical progress has been made by using constraint
programming \citep{DBLP:conf/cp/NdiayeS11,DBLP:conf/cp/McCreeshNPS16}, by
reducing to the maximum clique problem
\citep{LeviG,DBLP:conf/cp/McCreeshNPS16}, and by adapting subgraph isomorphism
algorithms \citep{UpcomingAAAIPaper}. This paper introduces a new branch and
bound algorithm which exploits special properties of these problems to allow a
much faster exploration of the search space, whilst retaining the filtering and
bounding benefits of the constraint programming approach. We describe the
algorithm for the basic maximum common subgraph problem, and discuss how it may
be adapted to handle vertex labels, edge labels, and the requirement that the
found subgraph be connected. We then present an empirical study of the
algorithm, demonstrating that it improves the state of the art by more than an
order of magnitude on the unlabelled variant of the problem, and showing that
it can handle much larger instances than earlier constraint programming or
clique approaches due to lower memory usage.

@@

\begin{algorithm}[ht]
\DontPrintSemicolon
\nl $\FuncSty{expand}(\AlgVar{future},M)$ \;
\nl \Begin{
%\nl \lIf {$\AlgVar{future} = \emptyset$ \bf{and} $|M| > |\AlgVar{incumbent}|$}
\nl \lIf {$|M| > |\AlgVar{incumbent}|$}{$save(matching)$} \label{StoreIncumbent}
%\nl \lIf {$\AlgVar{future} = \emptyset$}{return}
\nl $bf \gets 0$ \label{SetBestFutureToZero} \;
\nl \lFor {$\langle P,T \rangle \in \AlgVar{future}$}{$\AlgVar{bf} \gets \AlgVar{bf} + \min(|P|,|T|)$}
\nl \lIf {$|M|  + \AlgVar{bf} \leq |\AlgVar{incumbent}|$}{return} \label{PruneSearch}
\bigskip
\nl $\langle P,T \rangle \gets \FuncSty{SelectLabelClass}(\AlgVar{future})$ \label{SelectClass} \;
\nl $v \gets \FuncSty{SelectVertex}(P)$ \label{SelectVertex} \;
\nl \For {$w \in T$ \label{WLoop}} {
\nl    $M \gets M \cup (v,w)$ \label{GrowM} \;
\nl    \bf{Let} $\AlgVar{future'} \gets \emptyset$ \label{NewFuture} \;
\nl    \For {$\langle P',T'\rangle \in future$ \label{InnerLoop}}{
\nl        \bf{Let} $P'' \gets P' \cap N(v,G) \setminus \{v\}$ \label{NewPWithEdge} \;
\nl        \bf{Let} $T'' \gets T' \cap N(w,H) \setminus \{w\}$ \;
\nl        \If {$P'' \neq \emptyset$ \bf{and} $T'' \neq \emptyset$}
        {$\AlgVar{future'} \gets \AlgVar{future'} + \langle P'' , T'' \rangle$ \label{AddToFutureWithEdge}}
\nl        \bf{Let} $P'' \gets P' \cap N(v,\overline{G}) \setminus \{v\}$ \label{NewPWithoutEdge}  \;
\nl        \bf{Let} $T'' \gets T' \cap N(w,\overline{H}) \setminus \{w\}$ \;
\nl        \If {$P'' \neq \emptyset$ \bf{and} $T'' \neq \emptyset$}
        {$\AlgVar{future'} \gets \AlgVar{future'} + \langle P'' , T'' \rangle$} \label{InnerLoopEnd}
       }
\nl   $\FuncSty{expand}(\AlgVar{future'},M)$ \label{ExpandWithV} \;
\nl   $M \gets M \setminus \{(v,w)\}$ \label{ShrinkM} \;
  }
\nl $P' \gets P \setminus \{v\}$ \label{RemoveV} \;
\nl $\AlgVar{future} \gets \AlgVar{future} \setminus \{\langle P,T \rangle\}$\;
\nl \lIf {$P' \neq \emptyset$} {$\AlgVar{future} \gets \AlgVar{future} \cup \{\langle P',T \rangle \}$}
\nl $\FuncSty{expand}(\AlgVar{future},M)$ \label{ExpandWithoutV} \;
}
\;
\nl $\FuncSty{McSplit}(G,H)$ \label{McSplitFun} \;
\nl \Begin{
\nl $\AlgVar{incumbent} \gets \emptyset$ \;
\nl $\FuncSty{expand}(\{\langle V(G),V(H) \rangle \},\emptyset)$ \label{FirstExpandCall} \;
\nl return $|\AlgVar{incubent}|$ \;
}
\caption{\McSplit\ algorithm for Maximum Common Subgraph}
\label{jtAlg}
\end{algorithm}

@@

\begin{figure}[ht]
    \centering
    \subfigure[][MCS unlabelled undirected instances] {
        \centering
        \includegraphics*{gen-graph-plain-cumulative.pdf}
        \label{figure:plain-cumulative}
    }

@@

\paragraph{Large subgraph isomorphism instances} We also ran the algorithms on
a set of 5725 larger instances used in recent studies of subgraph
isomorphism~\cite{thelionpaper} and maximum common
subgraph~\cite{UpcomingAAAIPaper}.  This benchmark set includes real-world
graphs and graphs generated using random models.  Pattern graphs range from 4
vertices to 900 with a median of 80; target graphs range from 10 vertices to
6671 with a median of 561.  Cumulative runtimes including the kdown algorithm
from~\cite{UpcomingAAAIPaper} which was the previous state of the art for this
benchmark, are in \cref{figure:sip-cumulative}.  This is a challenging set of
instances, and more than half of the instances cannot be solved within a
timeout of 1000 seconds by any of the solvers. Moreover the CP-FC algorithm and
the clique encoding run out of memory on many of the instances.  \McSplit\
is the strongest solver overall for this dataset.
